High-accuracy neutron diffusion calculations based on integral transport theory

### 2019-12-03 docs/EPJP2019/response/response.tex

We hope that the Editor and the reviewer find the revised manuscript satisfactory and recommend it for publication.

%--------------------------------------------------------------
% LETTERS ENDS HERE ******************************************
\closing{\mbox{Sincerely yours on behalf of all the co-authors,}}

%---------------------------------------------------------------------------------------	-
\newpage

\textbf{Reviewer \#1}
\\ \\
% --------------------------------------------------------------------------------------
\textit{Please re-read the whole paper to check the language and the punctuation. Some sentences are very concise and could be expanded a little for clarity.}
\\ \\
\revised{The manuscript underwent a thorough examination of the language and punctuation, and some sentences were elaborated for clarity.}
\\ \\
% --------------------------------------------------------------------------------------
\textit{I have a concern on which the Authors may want to comment. Formula (3) becomes problematic where the gradient of the flux is small, which is usually the zone where diffusion works better.}
\\ \\
\revised{This is correct. Ronen did not elaborate on this issue in his technical note from 2004. Theoretically, both the numerator and the denominator should vanish at the same location, since both are accurate expressions for the neutron current at this location (where the flux gradient is weak). We did not try to find the analytic limit of this expression for $D$, however we assume that this ratio between two expressions, which vanish at the same location, approaches some constant value.\\ \\ Numerically, Tomatis and Dall'Osso 2011 chose to deal with this potential singularity by using the well-known drift terms, which serve as a \emph{numerical feature}. The (more accurate) surface current $J_{tr}$, calculated by integral expression, is written as the sum of the surface current calculated by diffusion, $J_D$, plus a correction $\delta J$. Hence, when the current vanishes, the correction $\delta J$ vanishes. Moreover, the correction is written as in Eq. (21), avoiding the potential divergence resulting from division by zero.}
\\ \\
% --------------------------------------------------------------------------------------
\textit{I do not think it would be too hard work to report more results. For instance, I think it would be nice to see the effect of the physical size of the domain, to better detect and evidence the effect of the boundary. This would add a lot to the value of the paper.}
\\ \\
\revised{An entire benchmark was added to the results section. The chosen benchmark is a one-dimensional highly heterogeneous core representing a BWR core. The sensitivity of the method to the slab width, especially at the boundaries, will be studied in future work, along with additional effects emanate from the boundaries conditions.}
\\ \\
% --------------------------------------------------------------------------------------
\textit{In the abstract: I suggest to avoid saying that the results match "extremely well" the reference... while this is true for the eigenvalue, it may be questionable for the flux near the boundary, where effects should be more deeply investigated. Maybe "well" is already enough.}
\\ \\
\revised{Fixed.}
\\ \\

### 2019-12-03 docs/EPJP2019/Gross-EPJP-2019.tex
A (almost final) revised version of the ms.
Tex file is updated:
	modified:   Gross-EPJP-2019.tex

New figures were added:
	modified:   figures/RM_dD1_100it_8_22_728.pdf
	modified:   figures/RM_dD2_100it_8_22_728.pdf
	new file:   figures/RM_flx_100it_8_22_728.pdf
	new file:   figures/Sn_Diff_RM_Tomatis2011.pdf
	new file:   figures/Sn_Diff_RM_core_1.pdf
	new file:   figures/Sn_Diff_RM_core_2.pdf
	new file:   figures/Sn_Diff_RM_core_3.pdf
	new file:   figures/dD_Tomatis2011_400_250it.pdf
	new file:   figures/flx_err_Tomatis2011_400_250.pdf
	new file:   figures/flx_err_diff_C1.pdf
	new file:   fi

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for The European Physical Journal PLUS
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag / Societa` Italiana di Fisica  %%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{filecontents}{leer.eps}
%!PS-Adobe-2.0 EPSF-2.0
%%%CreationDate: Mon Jul 13 16:51:17 1992
%%%DocumentFonts: (atend)
%%%Pages: 0 1
%%%BoundingBox: 72 31 601 342
%%%EndComments
%
%gsave
%72 31 moveto
%72 342 lineto
%601 342 lineto
%601 31 lineto
%72 31 lineto
%showpage
%grestore
%%%Trailer
%%%DocumentFonts: Helvetica
%\end{filecontents}
%
\documentclass[epj]{svjour}
% Remove option referee for final version
%
% Remove any % below to load the required packages
%\usepackage{latexsym}
\usepackage{mathtools} % Loads amsmath
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage{xfrac}
\usepackage{graphicx}
%\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref} % for hyperrefs
%\usepackage{cleveref}
%\usepackage[authoryear]{natbib}
\usepackage{gensymb}  % provides macro \degree which works in text and math
\usepackage{color}
%\usepackage{enumitem} 
\usepackage{bm}
\usepackage{multirow}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{arrows}

@@

% for flow charts
% Define block styles
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\tikzstyle{discre} = [rectangle, draw, fill=white!20, 
text width=6em, text centered, minimum height=3em, node distance=3.5cm, line width=1pt]
\tikzstyle{decision} = [diamond, draw, fill=white!20, 
text width=3em, text badly centered, node distance=2cm, inner sep=1pt, line width=1pt]
\tikzstyle{block} = [rectangle, draw, fill=white!20, 
text width=6em, text centered, minimum height=3em, node distance=3.5cm, line width=1pt]
\tikzstyle{blockEq} = [rectangle, draw, fill=white!20, 
text width=8em, text centered, minimum height=3em, node distance=3.5cm, line width=0pt]
\tikzstyle{cloud} = [draw, ellipse,fill=white!20, node distance=4cm,
text centered, text width=3em, minimum height=2em, line width=1pt]
\tikzstyle{line} = [draw, -{latex'}, line width=1.0pt]

@@

%
\graphicspath{ {./figures/} }
% ---------------------------------------------------
\begin{document}
%
\title{High-accuracy neutron diffusion calculations based on integral transport theory}
%\subtitle{}
\author{
	Roy Gross\inst{1} \and 
	Daniele Tomatis\inst{2} \and 
	Erez Gilad\inst{1}\thanks{\email{gilade@bgu.ac.il} (corresponding author)}
}                     % Do not remove
%
%\offprints{}          % Insert a name or remove this line
%
\institute{
	The Unit of Nuclear Engineering, Ben-Gurion University of the Negev, 8410501 Beer-Sheva, Israel 
	\and
	CEA, DEN, Service d'{\'e}tudes des r\'eacteurs et de math\'ematiques appliqu\'ees (SERMA),
	Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France}
%
\authorrunning{Gross, Tomatis \& Gilad} 
\titlerunning{Integral transport correction to diffusion calculations}
\date{Received: date / Revised version: date}
% The correct dates will be entered by Springer
%
\abstract{
In this paper, the Ronen method is developed, implemented, and applied to resolve the neutron flux and the criticality eigenvalue \revised{in one-dimensional simple homogeneous problem and in more complex heterogeneous benchmarks}. The Ronen method is based on iterative calculations of the multigroup diffusion coefficients using a multigroup diffusion model and \revised{is} driven by the solutions of the integral transport equation. The \revised{spatially-dependent} diffusion constants are modified \revised{locally} in order to produce new estimates of the surface currents \revised{using an integral transport expression}. The diffusion solver employed in this study uses finite differences and the transport-corrected currents are introduced into the numerical scheme \revised{using} drift terms. The corrected solutions are compared against reference results obtained by a discrete ordinate code. Boundary conditions are discussed and proper approximations are introduced to conserve the particle balance. The results match well with the reference solutions, especially in the limit of fine meshes, but slow convergence of the scalar flux is reported.
% 
\PACS{
	{28.00.00}{Nuclear engineering and nuclear power studies} \and
	{28.20.Gd}{Neutron transport: diffusion and moderation} \and
	{05.60.-k}{Transport processes} \and
	{28.41.-i}{Fission reactors}
     } % end of PACS codes
  \keywords{
  	neutron diffusion -- integral transport -- non-linear transport correction
  }
} %end of abstract
%
\maketitle
%
% ---------------------------------------------------
\section{Introduction}
\label{sec:intro}

Neutron transport calculations on a full core scale can be a highly intensive computational task. High-fidelity core design optimization or transient analyses can quickly become computationally impractical when using transport methods~\cite{Kim-2019}. For example, to achieve a 1\% accuracy in the local flux/power estimation, an order of 10\tsup{11} histories is needed in a full-core Monte Carlo calculation~\cite{Martin-2012}. Another problem is the huge number of variables and parameters to be stored and used during computation, e.g., tallies, geometry, cross sections, and depletion data. A conservative estimation of the memory needed for reasonably accurate full-core Monte Carlo neutronic calculation is in the range of terabytes (TBs)~\cite{Martin-2012}. \revised{The growing demand for high-accuracy and high-precision full-core computations challenges not only today’s high-end computing systems but will also challenge the near future (e.g., Exascale) computers~\cite{Martin-2012,Smith-2011,Kim-2019}.}

@@

In this paper, the development, implementation, and qualification of the Ronen Method~\cite{Ronen-2004} in one-dimensional homogeneous \revised{and heterogeneous} plane geometry is reported. This method is implemented as a highly accurate multigroup neutron diffusion solver based on novel transport corrections. The main hypothesis underlying the Ronen method is based on iterative calculations of the multigroup diffusion coefficients driven by the solutions of the integral transport equation~\cite{Ronen-2004,Tomatis-2011}.

@@

Since these accurate estimates of the currents are based on a known flux distribution, it was also suggested to execute new diffusion calculations, thus updating iteratively the diffusion coefficients in the global calculation. 
\begin{equation}\label{eq:RM-it}
D^{(k+1)}\rE = -\frac{\lvert\jtr^{(k)}\rE\rvert}{\lvert\bnabla\phi^{(k)}\rE\rvert} \,\, ,
\end{equation}
where $k$ is the iteration index. \revised{The use of a tensor notation is needed for the diffusion coefficient in general multidimensional problems.}

@@

where $E_1(x)$ is a first order exponential integral~\cite{Abramowitz-1964}. Note that this expression does not include the contribution of uncollided neutrons \revised{originated from an incoming angular} flux at the boundaries.

@@

In what follows, a generalization of~\eq{eq:flux-slab-l-0-no-bndry-terms} is derived for heterogeneous medium with anisotropic scattering.  
%
% ---------------------------------------------------
\subsection{Integral expressions for the neutron flux \revised{and} current}
\label{sec:int-flux-J}

@@

The angular dependence of the source and the uncollided boundary flux terms is expanded in a series of orthogonal Legendre polynomials
\begin{eqnarray}\label{eq:Q-series}
\psi(0,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(0,E) \pl{\mu} 	
} \nonumber \\
\psi(a,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(a,E) \pl{\mu} 	
} \,\, .
\end{eqnarray}
The source cross sections are written explicitly as~\cite{Tomatis-2011}
\begin{equation}\label{eq:src-mmnts}
\sigma_l(x,E\leftarrow E')dE' = 
\sigma_{s,l}(x,E\leftarrow E')dE'
+ \delta_{l0}\frac{\chi(E)}{\keff}\nu\sigma_f(x,E')dE' \,\, ,
\end{equation}
\revised{where
\begin{equation}\label{eq:src-mmnts-explicit}
\sigma_{s,l}(x,E\leftarrow E') = \int_{-1}^1 d\mu' P_l(\mu')\sigma_s(x,E\leftarrow E',\mu_0)
\end{equation}
with $\mu_0=\OO'\cdot\OO$} and the source term is expanded 
\begin{eqnarray}\label{eq:src}
Q(x,E,\mu)&=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_{s,l}(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} 
+ \frac{\chi(E)}{2\keff} \int_0^\infty { dE'
	\nu\sigma_f(x,E') \psi_0(x,E')
}  \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_l(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
\frac{2l+1}{2} \int_0^\infty { dE'
	q_l(x,E\leftarrow E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} q_l(x,E)\pl{\mu} 			
} 
\,\, ,
\end{eqnarray}
where the source angular moments are defined as
\begin{equation}\label{eq:q_l}
q_l(x,E) \equiv \int_0^\infty dE' q_l(x,E\leftarrow E') \equiv 
\int_0^\infty dE' \sigma_l(x,E\leftarrow E')\psi_l(x,E') dE' \,\, .
\end{equation}

@@

Using the definitions in~\eqsand{eq:JD-dJ}{eq:Ds} and~\fig{fig:mesh1D}, the discretized form of~\eq{eq:diff1} is
\begin{eqnarray}\label{eq:FD_diff_full}
- 2D_{i+1/2}^g\frac{\phi_{i+1,g}-\phi_{i,g}}{\dx_{i+1} + \dx_i}
- 2\delta D_{i+1/2}^g\frac{\phi_{i+1,g}+\phi_{i,g}}{(\dx_{i+1}+\dx_i)}
&+& 2D_{i-1/2}^g\frac{\phi_{i,g}-\phi_{i-1,g}}{\dx_{i} + \dx_{i-1}}  
+ 2\delta D_{i-1/2}^g\frac{\phi_{i,g}+\phi_{i-1,g}}{(\dx_i+\dx_{i-1})} 
\nonumber \\
&+& \sigma_{i}^g \phi_{i,g} 
= \sum\limits_{g'=1}^G \sigma_{s,i}^{g\leftarrow g'}\phi_{i,g'}
+ \frac{\chi_g}{\keff}\sum\limits_{g'=1}^G
 \nu\sigma_{f,i}^{g'}\phi_{i,g'} 
\,\,.
\end{eqnarray}
Rearranging \revised{terms, one gets}
\begin{eqnarray}\label{eq:FD_diff_org}
\left(
\frac{-D_{i-1/2}^g + \delta D_{i-1/2}^g }{\dx_{i-1}+\dx_{i}}
\right)
\phi_{i-1,g}
&+&\left[
\frac{D_{i+1/2}^g - \delta D_{i+1/2}^g}{\dx_{i+1}+\dx_i}
+\frac{D_{i-1/2}^g + \delta D_{i-1/2}^g}{\dx_{i}+\dx_{i-1}}
+\frac{\sigma_{i}^g}{2}
\right]
\phi_{i,g}
\nonumber \\
&+&\left(
\frac{-D_{i+1/2}^g -\delta D_{i+1/2}^g }{\dx_{i+1}+\dx_{i}}
\right)
\phi_{i+1,g} =
\frac{1}{2}q_{i,g}\,\, ,
\end{eqnarray}
where $q_{i,g}$ is the RHS of~\eq{eq:FD_diff_full}.

These equations can be formulated in operators notation according to
\begin{equation}\label{eq:op-not}
\mathcal{M}\Phi=\frac{1}{\keff}\mathcal{F}\Phi \,\, ,
\end{equation}
where $\mathcal{M}$ is the migration operator, whose entries are given by~\eq{eq:FD_diff_org} making it a three-diagonal banded matrix, and $\mathcal{F}$ is the neutron generation operator given by \revised{the second term in} $q_{i,g}$ \revised{(see Eqs.~\ref{eq:FD_diff_full}-\ref{eq:FD_diff_org})}. Note that $\mathcal{M}$ can be written as $\mathcal{M}=\mathcal{M}_0+\delta\mathcal{M}$ distinguishing the diffusion corrections from the constant diffusion coefficients derived from the transport problem. In the homogeneous case, $\mathcal{M}(x)=\mathcal{M}_0+\delta\mathcal{M}(x)$.
%
%Eq.~\eqref{eq:op-not} is solved using the power iteration algorithm, where the within-group equation is noted as $\mathcal{A}_g(x)\Phi_g(x)=\frac{1}{\keff}Q_g(x)$, where $\mathcal{A}_g(x)=\mathcal{A}_{g,0}+\delta\mathcal{A}_g(x)$.
%
% ---------------------------------------------------
\subsection{\revised{Evaluation of the currents by the integral form of the transport equation}}
\label{sec:calc-int-curr}

@@

	\draw[>=triangle 45, <->] (7cm,-2.1cm) -- (13cm,-2.1cm);
	\node[anchor=south] at (10cm,-2.1cm) {$\Xi_{i,j}^g + \sigma_{g,j}\left(x'-x_{j-1/2}\right)$};
	\end{tikzpicture}
	\bigbreak
	\caption{\revised{Notation for numerical evaluation of interface currents by the integral form of the transport equation.}}
	\label{fig:Xi}
\end{figure}

@@

%
% ---------------------------------------------------
\subsection{Boundary conditions}
\label{sec:BC}
A generalized form for the boundary condition (at the left) follows as
\begin{equation}\label{eq:BC}
J_D(x=0) = \revised{-D_0}\frac{\phi_0}{\Delta x_0/2 + \zeta} \,\, ,
\end{equation} 
where $\zeta$ is the extrapolation length in case of vacuum. \revised{The Marshak boundary conditions are reproduced by $\zeta=2D$, but the more accurate value $\zeta\approx 2.13D$ is
taken to match transport with better agreement~\cite{Lamarsh-1966}}. Reflection can be reproduced by $\zeta\rightarrow\infty$, whereas the condition of zero-flux is realized by $\zeta=0$. 
%
The quantity $\delta J$ at the boundary takes the simpler form of $\delta J = -\delta D_{-1/2}\phi_0$, without dividing by the spatial width, since no particular extrapolation length is appropriate for the correction. The expression for the right boundary is straightforward, implying a non-negative current. \revised{One should recall that the diffusion coefficient remains constant throughout the calculation, and so is the extrapolated distance. Only $\delta D$ is recalculated and the correction of the diffusion coefficient is implemented through the currents, i.e., $J_{tr} = J_D + \delta J$.}

@@

\revised{Two one-dimensional test cases are considered, homogeneous and heterogeneous. The reference solutions are calculated using a Python3 discrete ordinates ($S_N$) code with $N=16$~\cite{Lewis-1984}. The diffusion solutions and the Ronen iterations are produced using original codes developed in Pyhton3. All thresholds are set to 1E-6 and no (over/under)-relaxation is done. The initial diffusion coefficient is set to $D=(3\sigma_{tr})^{-1}=(3\sigma)^{-1}$. Void boundary conditions are used (extrapolated distance) at the edges of the slab and they are constant for each energy group.}

\subsection{Homogeneous case}
\label{sec:homog}

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for The European Physical Journal PLUS
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag / Societa` Italiana di Fisica  %%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{filecontents}{leer.eps}
%!PS-Adobe-2.0 EPSF-2.0
%%%CreationDate: Mon Jul 13 16:51:17 1992
%%%DocumentFonts: (atend)
%%%Pages: 0 1
%%%BoundingBox: 72 31 601 342
%%%EndComments
%
%gsave
%72 31 moveto
%72 342 lineto
%601 342 lineto
%601 31 lineto
%72 31 lineto
%showpage
%grestore
%%%Trailer
%%%DocumentFonts: Helvetica
%\end{filecontents}
%
\documentclass[epj]{svjour}
% Remove option referee for final version
%
% Remove any % below to load the required packages
%\usepackage{latexsym}
\usepackage{mathtools} % Loads amsmath
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage{xfrac}
\usepackage{graphicx}
%\usepackage[colorlinks=true,linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref} % for hyperrefs
%\usepackage{cleveref}
%\usepackage[authoryear]{natbib}
\usepackage{gensymb}  % provides macro \degree which works in text and math
\usepackage{color}
%\usepackage{enumitem} 
\usepackage{bm}
\usepackage{multirow}
\usepackage{cite}
\usepackage{subcaption} % For subfigure
\captionsetup{compatibility=false}
\usepackage{wrapfig}
%\usepackage{caption} % Caption in subfigure
\usepackage{tikz}
\usetikzlibrary{arrows}

@@

% for flow charts
% Define block styles

\usetikzlibrary{patterns}
\usetikzlibrary{shapes,arrows}
\tikzstyle{discre} = [rectangle, draw, fill=white!20, 
text width=6em, text centered, minimum height=3em, node distance=3.5cm, line width=1pt]
\tikzstyle{decision} = [diamond, draw, fill=white!20, 
text width=3em, text badly centered, node distance=2cm, inner sep=1pt, line width=1pt]
\tikzstyle{block} = [rectangle, draw, fill=white!20, 
text width=6em, text centered, minimum height=3em, node distance=3.5cm, line width=1pt]
\tikzstyle{blockEq} = [rectangle, draw, fill=white!20, 
text width=8em, text centered, minimum height=3em, node distance=3.5cm, line width=0pt]
\tikzstyle{cloud} = [draw, ellipse,fill=white!20, node distance=4cm,
text centered, text width=3em, minimum height=2em, line width=1pt]
\tikzstyle{line} = [draw, -{latex'}, line width=1.0pt]

@@

%
\graphicspath{ {./figures/} }
% ---------------------------------------------------
\begin{document}
%
\title{High-accuracy neutron diffusion calculations based on integral transport theory}
%\subtitle{}
\author{
	Roy Gross\inst{1} \and 
	Daniele Tomatis\inst{2} \and 
	Erez Gilad\inst{1}\thanks{\email{gilade@bgu.ac.il} (corresponding author)}
}                     % Do not remove
%
%\offprints{}          % Insert a name or remove this line
%
\institute{
	The Unit of Nuclear Engineering, Ben-Gurion University of the Negev, 8410501 Beer-Sheva, Israel 
	\and
	CEA, DEN, Service d'{\'e}tudes des r\'eacteurs et de math\'ematiques appliqu\'ees (SERMA),
	Universit\'e Paris-Saclay, F-91191 Gif-sur-Yvette, France}
%
\authorrunning{Gross, Tomatis \& Gilad} 
\titlerunning{Integral transport correction to diffusion calculations}
\date{Received: date / Revised version: date}
% The correct dates will be entered by Springer
%
\abstract{
In this paper, the Ronen method is developed, implemented, and applied to resolve the neutron flux and the criticality eigenvalue in one-dimensional simple homogeneous problem and in more complex heterogeneous benchmarks. The Ronen method is based on iterative calculations of the multigroup diffusion coefficients using a multigroup diffusion model and is driven by the solutions of the integral transport equation. The spatially-dependent diffusion constants are modified locally in order to produce new estimates of the surface currents using an integral transport expression. The diffusion solver employed in this study uses finite differences and the transport-corrected currents are introduced into the numerical scheme using drift terms. The corrected solutions are compared against reference results obtained by a discrete ordinate code. Boundary conditions are discussed and proper approximations are introduced to conserve the particle balance. The results match well with the reference solutions, especially in the limit of fine meshes, but slow convergence of the scalar flux is reported.
% 
\PACS{
	{28.00.00}{Nuclear engineering and nuclear power studies} \and
	{28.20.Gd}{Neutron transport: diffusion and moderation} \and
	{05.60.-k}{Transport processes} \and
	{28.41.-i}{Fission reactors}
     } % end of PACS codes
  \keywords{
  	neutron diffusion -- integral transport -- non-linear transport correction
  }
} %end of abstract
%
\maketitle
%
% ---------------------------------------------------
\section{Introduction}
\label{sec:intro}

Neutron transport calculations on a full core scale can be a highly intensive computational task. High-fidelity core design optimization or transient analyses can quickly become computationally impractical when using transport methods~\cite{Kim-2019}. For example, to achieve a 1\% accuracy in the local flux/power estimation, an order of 10\tsup{11} histories is needed in a full-core Monte Carlo calculation~\cite{Martin-2012}. Another problem is the huge number of variables and parameters to be stored and used during computation, e.g., tallies, geometry, cross sections, and depletion data. A conservative estimation of the memory needed for reasonably accurate full-core Monte Carlo neutronic calculation is in the range of terabytes (TBs)~\cite{Martin-2012}. The growing demand for high-accuracy and high-precision full-core computations challenges not only today’s high-end computing systems but will also challenge the near future (e.g., Exascale) computers~\cite{Martin-2012,Smith-2011,Kim-2019}.

@@

In this paper, the development, implementation, and qualification of the Ronen Method~\cite{Ronen-2004} in one-dimensional homogeneous and heterogeneous plane geometry is reported. This method is implemented as a highly accurate multigroup neutron diffusion solver based on novel transport corrections. The main hypothesis underlying the Ronen method is based on iterative calculations of the multigroup diffusion coefficients driven by the solutions of the integral transport equation~\cite{Ronen-2004,Tomatis-2011}.

@@

Since these accurate estimates of the currents are based on a known flux distribution, it was also suggested to execute new diffusion calculations, thus updating iteratively the diffusion coefficients in the global calculation. 
\begin{equation}\label{eq:RM-it}
D^{(k+1)}\rE = -\frac{\lvert\jtr^{(k)}\rE\rvert}{\lvert\bnabla\phi^{(k)}\rE\rvert} \,\, ,
\end{equation}
where $k$ is the iteration index. The use of a tensor notation is needed for the diffusion coefficient in general multidimensional problems.

@@

where $E_1(x)$ is a first order exponential integral~\cite{Abramowitz-1964}. Note that this expression does not include the contribution of uncollided neutrons originated from an incoming angular flux at the boundaries.

@@

In what follows, a generalization of~\eq{eq:flux-slab-l-0-no-bndry-terms} is derived for heterogeneous medium with anisotropic scattering.  
%
% ---------------------------------------------------
\subsection{Integral expressions for the neutron flux and current}
\label{sec:int-flux-J}

@@

The angular dependence of the source and the uncollided boundary flux terms is expanded in a series of orthogonal Legendre polynomials
\begin{eqnarray}\label{eq:Q-series}
\psi(0,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(0,E) \pl{\mu} 	
} \nonumber \\
\psi(a,E,\mu) &=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \psi_l(a,E) \pl{\mu} 	
} \,\, .
\end{eqnarray}
The source cross sections are written explicitly as~\cite{Tomatis-2011}
\begin{equation}\label{eq:src-mmnts}
\sigma_l(x,E\leftarrow E')dE' = 
\sigma_{s,l}(x,E\leftarrow E')dE'
+ \delta_{l0}\frac{\chi(E)}{\keff}\nu\sigma_f(x,E')dE' \,\, ,
\end{equation}
where
\begin{equation}\label{eq:src-mmnts-explicit}
\sigma_{s,l}(x,E\leftarrow E') = \int_{-1}^1 d\mu' P_l(\mu')\sigma_s(x,E\leftarrow E',\mu_0)
\end{equation}
with $\mu_0=\OO'\cdot\OO$ and the source term is expanded 
\begin{eqnarray}\label{eq:src}
Q(x,E,\mu)&=& \sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_{s,l}(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} 
+ \frac{\chi(E)}{2\keff} \int_0^\infty { dE'
	\nu\sigma_f(x,E') \psi_0(x,E')
}  \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} \int_0^\infty { dE'
		\sigma_l(x,E\leftarrow E')\psi_l(x,E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
\frac{2l+1}{2} \int_0^\infty { dE'
	q_l(x,E\leftarrow E')\pl{\mu} 			
	}
} \nonumber \\
&=&\sum\limits_{l=0}^\infty {
	\frac{2l+1}{2} q_l(x,E)\pl{\mu} 			
} 
\,\, ,
\end{eqnarray}
where the source angular moments are defined as
\begin{equation}\label{eq:q_l}
q_l(x,E) \equiv \int_0^\infty dE' q_l(x,E\leftarrow E') \equiv 
\int_0^\infty dE' \sigma_l(x,E\leftarrow E')\psi_l(x,E') dE' \,\, .
\end{equation}

@@

Using the definitions in~\eqsand{eq:JD-dJ}{eq:Ds} and~\fig{fig:mesh1D}, the discretized form of~\eq{eq:diff1} is
\begin{eqnarray}\label{eq:FD_diff_full}
- 2D_{i+1/2}^g\frac{\phi_{i+1,g}-\phi_{i,g}}{\dx_{i+1} + \dx_i}
- 2\delta D_{i+1/2}^g\frac{\phi_{i+1,g}+\phi_{i,g}}{(\dx_{i+1}+\dx_i)}
&+& 2D_{i-1/2}^g\frac{\phi_{i,g}-\phi_{i-1,g}}{\dx_{i} + \dx_{i-1}}  
+ 2\delta D_{i-1/2}^g\frac{\phi_{i,g}+\phi_{i-1,g}}{(\dx_i+\dx_{i-1})} 
\nonumber \\
&+& \sigma_{i}^g \phi_{i,g} 
= \sum\limits_{g'=1}^G \sigma_{s,i}^{g\leftarrow g'}\phi_{i,g'}
+ \frac{\chi_g}{\keff}\sum\limits_{g'=1}^G
 \nu\sigma_{f,i}^{g'}\phi_{i,g'} 
\,\,.
\end{eqnarray}
Rearranging terms, one gets
\begin{eqnarray}\label{eq:FD_diff_org}
\left(
\frac{-D_{i-1/2}^g + \delta D_{i-1/2}^g }{\dx_{i-1}+\dx_{i}}
\right)
\phi_{i-1,g}
&+&\left[
\frac{D_{i+1/2}^g - \delta D_{i+1/2}^g}{\dx_{i+1}+\dx_i}
+\frac{D_{i-1/2}^g + \delta D_{i-1/2}^g}{\dx_{i}+\dx_{i-1}}
+\frac{\sigma_{i}^g}{2}
\right]
\phi_{i,g}
\nonumber \\
&+&\left(
\frac{-D_{i+1/2}^g -\delta D_{i+1/2}^g }{\dx_{i+1}+\dx_{i}}
\right)
\phi_{i+1,g} =
\frac{1}{2}q_{i,g}\,\, ,
\end{eqnarray}
where $q_{i,g}$ is the RHS of~\eq{eq:FD_diff_full}.

These equations can be formulated in operators notation according to
\begin{equation}\label{eq:op-not}
\mathcal{M}\Phi=\frac{1}{\keff}\mathcal{F}\Phi \,\, ,
\end{equation}
where $\mathcal{M}$ is the migration operator, whose entries are given by~\eq{eq:FD_diff_org} making it a three-diagonal banded matrix, and $\mathcal{F}$ is the neutron generation operator given by the second term in $q_{i,g}$ (see Eqs.~\ref{eq:FD_diff_full}-\ref{eq:FD_diff_org}). Note that $\mathcal{M}$ can be written as $\mathcal{M}=\mathcal{M}_0+\delta\mathcal{M}$ distinguishing the diffusion corrections from the constant diffusion coefficients derived from the transport problem. In the homogeneous case, $\mathcal{M}(x)=\mathcal{M}_0+\delta\mathcal{M}(x)$.
%
%Eq.~\eqref{eq:op-not} is solved using the power iteration algorithm, where the within-group equation is noted as $\mathcal{A}_g(x)\Phi_g(x)=\frac{1}{\keff}Q_g(x)$, where $\mathcal{A}_g(x)=\mathcal{A}_{g,0}+\delta\mathcal{A}_g(x)$.
%
% ---------------------------------------------------
\subsection{Evaluation of the currents by the integral form of the transport equation}
\label{sec:calc-int-curr}

@@

	\draw[>=triangle 45, <->] (7cm,-2.1cm) -- (13cm,-2.1cm);
	\node[anchor=south] at (10cm,-2.1cm) {$\Xi_{i,j}^g + \sigma_{g,j}\left(x'-x_{j-1/2}\right)$};
	\end{tikzpicture}
	\bigbreak
	\caption{Notation for numerical evaluation of interface currents by the integral form of the transport equation.}
	\label{fig:Xi}
\end{figure}

@@

%
% ---------------------------------------------------
\subsection{Boundary conditions}
\label{sec:BC}
A generalized form for the boundary condition (at the left) follows as
\begin{equation}\label{eq:BC}
J_D(x=0) = -D_0\frac{\phi_0}{\Delta x_0/2 + \zeta} \,\, ,
\end{equation} 
where $\zeta$ is the extrapolation length in case of vacuum. The Marshak boundary conditions are reproduced by $\zeta=2D$, but the more accurate value $\zeta\approx 2.13D$ is
taken to match transport with better agreement~\cite{Lamarsh-1966}. Reflection can be reproduced by $\zeta\rightarrow\infty$, whereas the condition of zero-flux is realized by $\zeta=0$. 
%
The quantity $\delta J$ at the boundary takes the simpler form of $\delta J = -\delta D_{-1/2}\phi_0$, without dividing by the spatial width, since no particular extrapolation length is appropriate for the correction. The expression for the right boundary is straightforward, implying a non-negative current. One should recall that the diffusion coefficient remains constant throughout the calculation, and so is the extrapolated distance. Only $\delta D$ is recalculated and the correction of the diffusion coefficient is implemented through the currents, i.e., $J_{tr} = J_D + \delta J$.

@@

Two one-dimensional test cases are considered, homogeneous and heterogeneous. The reference solutions are calculated using a Python3 discrete ordinates ($S_N$) code with $N=16$~\cite{Lewis-1984}. The diffusion solutions and the Ronen iterations are produced using original codes developed in Pyhton3. All thresholds are set to 1E-6 and no (over/under)-relaxation is done. The initial diffusion coefficient is set to $D=(3\sigma_{tr})^{-1}=(3\sigma)^{-1}$. Void boundary conditions are used (extrapolated distance) at the edges of the slab and they are constant for each energy group.

%
% ---------------------------------------------------
\subsection{Homogeneous case}
\label{sec:homog}
